The growth of children's semantic and phonological networks: insight from 10 languages

### 2019-02-12 journal/ms.Rmd
final version

 Children tend to produce words earlier when they are connected to a variety of other words along the phonological and semantic dimensions. Though this connectivity effect has been extensively documented, little is known about the underlying developmental mechanism.  One view suggests that learning is primarily predicted by a network growth model where highly connected words in the child's early lexicon enable learning of similar words. Another view suggests that learning is driven by highly connected words in the external learning environment, instead of highly connected words in the early internal lexicon. The present study tests both scenarios systematically in both the phonological and semantic domains across 10 languages. We show that external connectivity in the learning environment drives growth in both production- and comprehension-based vocabularies, even controlling for word frequency and length. This pattern of findings suggests a word learning mechanism where children harness their statistical learning abilities to detect and learn highly connected words in the learning environment.

@@

What factors shape vocabulary learning over the course of early childhood? To investigate this question, scientists have adopted multiple research strategies, from conducting controlled laboratory experiments [e.g. @markman90] to analyzing dense corpora capturing language learning in context [e.g., @roy2015]. One prominent strategy consists in documenting the timeline of words' acquisition, and studying the properties that make words easy or hard to learn [e.g., @goodman2008; @huttenlocher1991]. For example, @goodman2008 found that, within a lexical category (e.g., nouns), higher parental frequency is associated with earlier learning. Researchers have studied the role of several other factors such as word length and the mean length of utterances in which the word occurs [e.g., @braginsky2016; @swingley2018].

Besides word-level properties, the structure of the lexicon (that is, how words relate to one another) was also found to predict the Age of Acquisition (AoA) of words. The lexical structure can be characterized in terms of a network where each node represents a word in the vocabulary, and each link between two nodes represents a relationship between the corresponding pair of words [e.g., @collins1975; @luce1998]. Previous studies have investigated early vocabulary structure by constructing networks using  a variety of word-word relations including shared semantic features [@mcrae2005], target-cue relationships in free association norms [@nelson1998], co-occurrence in child directed speech [@macwhinney2014], and phonological relatedness [@vitevitch2008]. These studies have generally found that children tend to produce nouns that have higher neighborhood density (i.e., high connectivity in the network) earlier, both at the phonological and the semantic level [@storkel2009; @hills2009; @hills2010; @stella2017; @carlson2014].

@@

```{r growth, fig.cap = "Illustration of the two growth scenarios. Filled grey circles represent known words (Internal) at a certain point in time. The empty, numbered circles represent words that have not yet been learned (External) and which are candidates to enter the lexicon next. The identity of the word that is going to be learned depends on the growth scenario.  Here the squares indicate the node that drives growth in each scenario and the hand pointer indicates which word is likely to be learned. For INT, the utility of a candidate, external node is the average degree (i.e., number of links) of the internal nodes that it would attach to. Thus, according to INT, node 1 is more likely to enter the lexicon. For EXT, the utility of a candidate node is its degree in the entire network. According to EXT, node 2 is more likely to enter the lexicon next.", fig.align = "center", out.width = "300px"}
knitr::include_graphics("figs/growth3.png")
```

@hills2009 investigated the growth of lexico-semantic networks in toddlers and found that growth did not proceed according to INT as was originally hypothesized by @steyvers2005, but rather according to EXT.\footnote{Besides INT and EXT, the authors tested a third mechanism (called the lure of associates) which resembles EXT in that it is driven by the connectivity of external nodes, except that this connectivity is computed with respect to words that are known. However, EXT is the externally-driven scenario that best predicted the data in this previous work.} This is an important finding because it suggests that learning in the early stages is mostly driven by properties of the external input, regardless of how past knowledge is organized. However, this work has been done in the special case of networks that are based on 1) semantic associations, 2) production-based vocabularies, and 3) data from English-learning children, only. The extent to which this result depends on the domain (e.g., semantic vs. phonological connectivity), the vocabulary measure (production vs. comprehension) and culture/language is still unclear. In this work, we test the generality of the finding along these three dimensions.

First, we study the phonological network in addition to the semantic network. These two networks represent different ways the mental lexicon is structured [@beckage2016]. In particular, words that are neighbors in the semantic network (e.g., "cat", "dog") are not necessarily neighbors in the phonological network, and vice versa. Does the phonological network also predict word learning? Previous work did find an effect of nouns' connectivity in the phonological network on their age of learning [@storkel2009; @stella2017; @carlson2014]. In other words, nouns learned earlier in life tend to sound similar to many other nouns than a word learned later in life. However, this finding is *a priori* compatible with both INT and EXT, and previous studies did not explicitly compare these two mechanisms.  Here, we investigate whether phonological networks, like semantic networks, grow through EXT, or if they rather grow via INT (Figure\ \@ref(fig:growth)).
<!--
The phonological strcuture influence word processing in adults (Vitevitch, 2008;  Chan & Vitevitch, 2009; Chan & Vitevitch,2010; Vitevitch, Chan, and Roodenrys, 2012; Castro et al. 2018). Here we test whether the mechanism of networtk growth in the phonological domains mirrors the mechanism found in the semantic domains. -->

Second, we study vocabularies measured using both comprehension and production. Previous studies have found differences between these vocabularies in terms of their content and rate of acquisition [@benedict1979; @fenson94; @bates1995]. These differences may reflect the fact that comprehension and production do not share the same constraints. For instance, whereas comprehension depends on the ease with which words are stored and accessed, production depends, additionally, on the ease with which words are articulated, e.g., shorter words are produced earlier [@braginsky2016]. By investigating comprehension-based vocabularies, we assess the extent to which the network growth mechanism captures general learning patterns beyond the specific constraints of production.

Finally, we use developmental data in 10 languages. Lexical networks can show more or less cross-linguistic variability along both the semantic and phonological domains [@youn2016; @arbesman2010;@lupyan2017]. Besides, cultures might differ in the way caregivers talk to children [@kuhl1997; @cristia2017], and this difference in the input could influence the way in which the children's networks grow. Thus, cross-linguistic comparison is crucial to test what growth mechanism is cognitively universal and is used by all children, and what mechanism is specific to some patterns of learning that emerge due to the particulars of a given language or culture [@bates1987; @slobin2014]. 
 <!--
\footnote{The difference is more obvious in the phonological case where the network structure––and the distribution of word connectivity––can a priori change from language to language depending on various languistic conventions.}
-->

We adopted the following research strategy. We used a cross-linguistic dataset of children's vocabulary development [@frank2017; @fenson94], and we studied the timeline of noun learning using the normative age of acquisition (i.e., the age at which at least 50% of children know a given words). Our choice of studying the normative learning trajectory rather than the individual trajectories was motivated by the nature of the dataset used---which is primarily based on cross-sectional studies. Children may vary in their individual learning trajectories, but the aggregate data provide highly robust measures of the average learning patterns [@fenson94]. The use of such measures has lead to important insights on the mechanisms of word learning [@storkel2009; @stella2017; @hills2009; @hills2010; @goodman2008; @braginsky2016].

The paper is organized as follows. First, we describe the datasets we used and explain how we constructed the networks. Second, we analyze static properties of words' connectivity in these networks (correlation with age of acquisition and shape of the distribution) and we explain how these properties inform hypotheses about network growth. Next, we explicitly fit the two hypothesized growth mechanisms to the data. We investigate the extent to which the results obtained in @hills2009 generalize to phonological networks and comprehension-based vocabularies, and whether this generalization holds cross-linguistically.

@@

## Age of acquisition
For each word in the CDI data, we compute the proportion of children who understand or produce the word at each month. Then we fit a logistic curve to these proportions and determined when the curve crosses 0.5, i.e., the age at which at least 50% of children know the word. We take this point in time to be each word’s age of acquisition [@goodman2008; @braginsky2016].
<!--We obtained these nouns in ten languages: Croatian, Danish, English, French, Italian, Norwegian, Russian, Spanish, Swedish, and Turkish. We used the subset of nouns that had entries in the Florida Association Norms (see below). Since these norms are available only in English, we used the hand-checked translation equivalents provided by @braginsky2016, allowing us to use the English association norms across languages. Table \@ref(tab:stats) gives an overview of the data used. -->

@@

kable(statistics, format = "latex", escape = FALSE, booktabs = TRUE,
      linesep = "", format.args = list(big.mark = ","),
      caption = "Statistics for the dataset we used.",
      col.names = c("Language",  "Nouns", "Ages", "Nouns", "Ages")) %>%
  add_header_above(c("", "Comprehension" = 2, "Production" = 2)) %>%
  column_spec(1, bold = TRUE) %>%
  kable_styling(position = "center")

@@

Since  the free association norms are available only in English, we used the hand-checked translation equivalents available in Wordbank, which allowed us to use the English association norms across 10 languages. Using the same association data across languages does not necessarily lead to similar networks. Indeed, though this approximation assumes that the semantic similarity measure is universal---which is a reasonable approximation [e.g., @youn2016]---the set of words acquired by children as well as the timeline of this acquisition can still vary from language to language leading to possibility different learning strategies.

@@

In previous studies, two nodes were linked if they had an edit distance of 1 [@storkel2009; @carlson2014; @stella2017]. However, these studies reported a contribution of phonological connectivity to noun learning when networks were built using a rich adult vocabulary. Since the focus of the current study is on the mechanism of growth, the networks should be based on children's early vocabulary. The latter, however, contains very few noun pairs with an edit distance of 1. Thus, we increased the threshold from 1 to 2, that is, two nodes were related if their edit distance was equal to 1 or 2.\footnote{We also considered the case of an edit distance of 1 as well as the continuous measure, i.e., the inverse edit distance without threshold. In both cases, the results were weaker than those obtained with a threshold of 2. We did not consider the case of a threshold larger than 2 since many pairs become phonologically unrelated.} The connectivity of a given node was characterized with its *degree*: the number of links it shares with other words.

@@

We start by analyzing word connectivity in the global (static) network. We constructed this network using nouns learned by the oldest age for which we have CDI data (e.g., in English this corresponds, in comprehension, to the network by 18 months, and in production, to the network by 30 months). This global network is the end-state towards which both INT and EXT converge by the last month of learning. Moreover, following @hills2009, we used this end-state network as a proxy for the external connectivity in the learning environment. Below we analyze properties of this global networks that may a priori hint at an INT- or EXT-like growth.

@@

```{r corrProd, out.width = "\\textwidth", fig.pos = "!h", fig.height=5, fig.cap = "Age of production in the global network as predicted by the degree (i.e., connectivity) in this network. Results are shown in each language for phonological and semantic networks. Each point is a word, with lines indicating linear model fits, and numbers indicating the Pearson correlation coefficients."}

@@

```{r corrComp, out.width = "\\textwidth", fig.pos = "!h", fig.height=5, fig.cap = "Age of comprehension in the global network as predicted by the degree (i.e., connectivity) in this network. Results are shown in each language for phonological and semantic networks. Each point is a word, with lines indicating linear model fits, and numbers indicating the Pearson correlation coefficients."}
corr_comp
#plot_grid(plot_correlation_prod , NULL, plot_correlation_comp , labels = c("A", "", "B"), ncol = 1, align = "v", rel_heights = c(1.3, 0.1, 1.3))

@@

Figures \@ref(fig:corrProd) and \@ref(fig:corrComp) show how the age of production and comprehension, respectively, correlates with the degree (or indegree for the semantic networks). For ease of visual comparison, the predictor (i.e., the degree) was centered and scaled. The plots show, overall, a negative correlation between the month of acquisition and the degree.  In production data, the average correlation across languages was `r ps_ave` ($SD=$ `r ps_sd`) for the semantic networks and `r pp_ave` ($SD=$ `r pp_sd`) for the phonological networks.  In comprehension data, the average correlation was `r cs_ave` ($SD=$ `r cs_sd`) for the semantic networks and `r cp_ave` ($SD=$ `r cp_sd`) for the phonological networks. These results indicate that nouns with higher degrees are generally learned earlier, thus replicating previous findings in English [@storkel2009; @hills2009] and extending these findings to 10 different languages, generally, in both production- and comprehension-based vocabularies.

@@

We estimated the parameter $\beta$ using a Bayesian approach. The inference was performed using the probabilistic programming language  WebPPL [@dippl]. We defined a uniform prior over $\beta$, and at each month, we computed the likelihood function over words that could possibly enter the lexicon at that month, fit to the words that have been learned at that month (using formula 1). Markov Chain Monte Carlo sampling resulted in a posterior distribution over $\beta$, which we summarized in Figure \@ref(fig:growthPred). The results replicate Hills et al.'s original finding regarding the semantic network in English and the production-based vocabulary, which is that this network grows by EXT, not by INT. Crucially, our results show that, generally speaking, this finding generalizes to comprehension-based vocabulary, and holds across languages. This generalization was obtained in both the semantic^[One could imagine that the fact of using English free association norms cross-linguistically would decrease the effect of non-English semantic networks because of possible cultural differences. However, our findings do not support this assumption, rather, it supports our initial approximation about the universality of the semantic similarity measure. That said, this approximation is not perfect. For example there is evidence that a small part of the variance in free association data can be explained by phonological similarity [@kachergis2011; @matusevych2018], thus leading to possibly minor cross-linguistic differences.] and phonological domains.

@@

Above we showed that the way semantic and phonological information is structured in the learning environment contributes to noun learning (via EXT) across languages. However, we know that other factors influence learning as well [e.g., @braginsky2016]. Next we investigated how semantic and phonological connectivity interact with two other factors. The first one is word frequency, a well studied factor shown to predict the age of acquisition in a reliable fashion [e.g. @goodman2008]. The second factor is word length, which was shown to correlate with phonological connectivity: Shorter words are more likely to have higher connectivity [@pisoni1985; @vitevitch2005].

@@

For word length, we counted the number of phonemes in our generated IPA transcription. For word frequency, we used the frequency estimates from @braginsky2016 where unigram counts were derived based on CHILDES corpora in each language [@macwhinney2014]. Although these frequency counts use transcripts from independent sets of children, they are based on large samples, and this allows us to average out possible differences between children and the specificities of their input [see @goodman2008 for a similar research strategy].

@@

The findings for the new predictors were as follows. Overall, frequency is the largest and most consistent predictor of age of acquisition in both comprehension and production data and across languages, endorsing results for nouns across a variety of analyses [@goodman2008; @braginsky2016; @roy2015]. Word length is more predictive for production than comprehension (and this difference is very clear in the global model), replicating previous work (Braginsky et al., under review). Thus, word length seems to reflect the effects of production's constraints rather than comprehension's constraints, i.e., longer words are harder to articulate but they may not be significantly more difficult to store and access.

@@

We found that the original findings reported in @hills2009 generalize well across all these dimensions. First, just like semantic networks, phonological networks grow via the externally-driven scenario (EXT), not by the internally-driven mechanism (INT). Second, comprehension-based vocabularies grow in a way similar to production-based vocabularies. Finally, the findings were, overall, similar across the 10 languages we tested. Although we find some cross-linguistic variation when semantic and phonological networks were pitted against frequency and length, this variability is to be taken with a grain of salt as it might be exaggerated in our study by the limited and partially-overlapping sample of nouns for each language. In fact, both phonological and semantic connectivity are significant predictors above and beyond frequency and length when data are pooled across languages.

@@

For example, in the semantic domain, growth according to EXT can be explained by a mechanism similar to cross-situational learning [@smith2008; @mcmurray2012; @yurovsky2015]. According to this mechanism, children track the co-occurrence of concrete nouns with their possible semantic referents. The referent of a word heard in only one naming situation can be ambiguous (e.g., when the word "ball" is heard for the first time in the presence of both a ball and a chair), but hearing the same word in a diversity of semantic contexts allows the learner to narrow down the set of possible word-object mappings. In our case, free association (used to determine semantic network connectivity) is related to contextual co-occurrence [@griffiths07; @fourtassi2013], meaning that highly connected words will tend to occur in a variety of speech and referential contexts.  This fact makes such words easier to learn because they have more referential disambiguating cues across learning contexts. Crucially, children can learn these words without necessarily knowing the meaning of all other words with witch they co-occur (hence the similarity with EXT). This possibility is supported by the finding that words' diversity of occurrence in child directed speech predicts their age of learning [@stella2017; @hills2010].

@@

This work shares a number of limitations with previous studies using similar research strategy and datasets. Chief among these limitations is the fact that the age of word acquisition is computed using different children at different ages (due to the fact that available CDI data is mainly cross-sectional). Such a measure has been shown to be valid and reliable [@fenson94], and has allowed researchers to study important aspects of word learning [@hills2009; @stella2017; @storkel2009; @goodman2008; @braginsky2016 ]. In our case, the use of cost-effective cross-sectional data has allowed us to leverage large-scale studies across several languages. That said, it is important to remember that this type of data can only inform us about the learning trajectory of the "average" child. Although our study endorses, overall, the externally-driven account of network growth, this does not mean individual children never use some variant of INT or some combination of both INT and EXT (Beckage and Colunga, under review). To illustrate, some children develop "islands of expertise", that is, well organized knowledge about a certain topic (e.g., birds or dinosaurs). This prior knowledge enables these children to learn new related words more easily [e.g., @chi1983].

>>>>

 Children tend to produce words earlier when they are connected to a variety of other words along the phonological and semantic dimensions. Though these semantic and phonological connectivity effects have been extensively documented, little is known about their underlying developmental mechanism. One possibility is that learning is driven by lexical network growth where highly connected words in the child's early lexicon enable learning of similar words. Another possibility is that learning is driven by highly connected words in the external learning environment, instead of highly connected words in the early internal lexicon. The present study tests both scenarios systematically in both the phonological and semantic domains across 10 languages. We show that phonological and semantic connectivity in the learning environment drives growth in both production- and comprehension-based vocabularies, even controlling for word frequency and length. This pattern of findings suggests a word learning process where children harness their statistical learning abilities to detect and learn highly connected words in the learning environment.

@@

What factors shape vocabulary learning over the course of early childhood? To investigate this question, scientists have adopted multiple research strategies, from conducting controlled laboratory experiments [e.g. @markman90] to analyzing dense corpora capturing language learning in context [e.g., @roy2015]. One prominent strategy consists in documenting the timeline of words' acquisition, and studying the properties that make words easy or hard to learn [e.g., @goodman2008; @huttenlocher1991]. For example, @goodman2008 found that, within a lexical category (e.g., nouns), higher parental frequency is associated with earlier learning. Researchers have studied the role of several other factors such as word length and the mean length of utterances in which the word occurs [e.g., @braginsky; @swingley2018].

Besides word-level properties, the structure of the lexicon (that is, how words relate to one another) is also linked to the the Age of Acquisition (AoA) of words. The lexical structure can be characterized in terms of a network where each node represents a word in the vocabulary, and each link between two nodes represents a relationship between the corresponding pair of words [e.g., @collins1975; @luce1998]. Previous studies have investigated early vocabulary structure by constructing networks using  a variety of word-word relations including shared semantic features [@mcrae2005], target-cue relationships in free association norms [@nelson1998], co-occurrence in child directed speech [@macwhinney2014], and phonological relatedness [@vitevitch2008]. These studies have generally found that children tend to produce words that have higher neighborhood density (i.e., high connectivity in the network) earlier, both at the phonological and the semantic level [@storkel2009; @hills2009; @hills2010; @stella2017; @carlson2014].

@@

```{r growth, fig.cap = "Illustration of the two growth scenarios. Filled grey circles represent known words (Internal) at a certain point in time. The empty, numbered circles represent words that have not yet been learned (External) and which are candidates to enter the lexicon next. The identity of the word that is going to be learned depends on the growth scenario.  Here the squares indicate the node that drives growth in each scenario and the hand pointer indicates which word is likely to be learned. For INT, the utility of a candidate, external node is the average degree (i.e., number of links) of the internal nodes that it would attach to. Thus, according to INT, node 1 is more likely to enter the lexicon. For EXT, the utility of a candidate node is its degree in the entire network. According to EXT, node 2 is more likely to enter the lexicon next. This figure is based on an example from Hills et al (2009).", fig.align = "center", out.width = "300px"}
knitr::include_graphics("figs/growth3.png")
```

@hills2009 investigated the growth of lexico-semantic networks in toddlers and found that growth did not proceed according to INT as was originally hypothesized by @steyvers2005, but rather according to EXT.\footnote{Besides INT and EXT, the authors tested a third mechanism (called the lure of associates) which resembles EXT in that it is driven by the connectivity of external nodes, except that this connectivity is computed with respect to words that are known. However, EXT is the externally-driven scenario that best predicted the data in this previous work.} This is an important finding because it suggests that learning in the early stages is mostly driven by properties of the external input, regardless of how past knowledge is organized. However, this work explored the INT/EXT growth in a special case: networks that were based on 1) semantic associations, 2) production-based vocabularies, and 3) data from English-learning children, only. The extent to which this result depends on the domain (e.g., semantic vs. phonological connectivity), the vocabulary measure (production vs. comprehension) and culture/language is thus an open area for investigation [@hills2018]. In this work, we test the generality of prior findings along these three dimensions.

First, we study the phonological network in addition to the semantic network. These two networks represent different ways the mental lexicon is structured [@beckage2016]. In particular, words that are neighbors in the semantic network (e.g., "cat", "dog") are not necessarily neighbors in the phonological network, and vice versa. Does the phonological network also predict word learning? Previous work has found an effect of words' connectivity in the phonological network on their age of learning [@storkel2009; @stella2017; @carlson2014]. In other words, words learned earlier in life tend to sound similar to many other words than a word learned later in life. However, this finding is *a priori* compatible with both INT and EXT, and previous studies did not explicitly compare these two mechanisms.  Here, we investigate whether phonological networks, like semantic networks, grow through EXT, or if they rather grow via INT (Figure\ \@ref(fig:growth)).
<!--
The phonological strcuture influence word processing in adults (Vitevitch, 2008;  Chan & Vitevitch, 2009; Chan & Vitevitch,2010; Vitevitch, Chan, and Roodenrys, 2012; Castro et al. 2018). Here we test whether the mechanism of networtk growth in the phonological domains mirrors the mechanism found in the semantic domains. -->

Second, we study vocabularies measured using both comprehension and production. Previous studies have found differences between these vocabularies in terms of their content and rate of acquisition [@benedict1979; @fenson94; @bates1995]. These differences may reflect the fact that comprehension and production do not share the same constraints. For instance, whereas comprehension depends on the ease with which words are stored and accessed, production depends, additionally, on the ease with which words are articulated, e.g., shorter words are produced earlier [@braginsky]. By investigating comprehension-based vocabularies, we assess the extent to which the network growth mechanism captures general learning patterns beyond the specific constraints of production.

Finally, we use developmental data in 10 languages. Lexical networks can show more or less cross-linguistic variability along both the semantic and phonological domains [@youn2016; @arbesman2010;@lupyan2017]. Besides, cultures might differ in the way caregivers talk to children [@kuhl1997; @cristia2017], and this difference in the input could influence the way in which the children's networks grow. Thus, cross-linguistic comparison is crucial to test the extent to which growth mechanisms are equally engaged across a wider variety of cultures compared with the extent to which the growth mechanisms are specific to patterns of learning that emerge due to the particulars of a given language or culture [@bates1987; @slobin2014]. 
 <!--
\footnote{The difference is more obvious in the phonological case where the network structure––and the distribution of word connectivity––can a priori change from language to language depending on various languistic conventions.}
-->

We adopted the following research strategy. We used parent reports on the MacArthur-Bates Communicative Development Inventory and its cross-linguistic adaptations [@frank2017; @fenson94]. We studied the timeline of word learning using the normative age of acquisition (i.e., the age at which at least 50% of children know a given words). Our choice of studying the normative learning trajectory rather than the individual trajectories was motivated by the nature of the dataset used---which is primarily based on cross-sectional studies. Children may vary in their individual learning trajectories, but the aggregate data provide highly robust measures of the average learning patterns [@fenson94]. The use of such measures has lead to important insights on the mechanisms of word learning [@storkel2009; @stella2017; @hills2009; @hills2010; @goodman2008].

The paper is organized as follows. First, we describe the datasets we used and explain how we constructed the networks. Second, we analyze static properties of words' connectivity in these networks (correlation with age of acquisition and shape of the distribution) and we explain how these properties inform hypotheses about network growth. Next, we fit the two hypothesized growth mechanisms to the data. We investigate the extent to which the results obtained in @hills2009 generalize to phonological networks and comprehension-based vocabularies, and whether this generalization holds cross-linguistically.

@@

## Age of acquisition
For each word in the CDI data, we compute the proportion of children who understand or produce the word at each month. Then we fit a logistic curve to these proportions and determined when the curve crosses 0.5, i.e., the age at which at least 50% of children know the word. We take this point in time to be each word’s age of acquisition [@goodman2008; @braginsky].
<!--We obtained these nouns in ten languages: Croatian, Danish, English, French, Italian, Norwegian, Russian, Spanish, Swedish, and Turkish. We used the subset of nouns that had entries in the Florida Association Norms (see below). Since these norms are available only in English, we used the hand-checked translation equivalents provided by @braginsky2016, allowing us to use the English association norms across languages. Table \@ref(tab:stats) gives an overview of the data used. -->

@@

kable(statistics, format = "latex", escape = FALSE, booktabs = TRUE,
      linesep = "", format.args = list(big.mark = ","),
      caption = "Statistics for the dataset we used.",
      col.names = c("Language",  "Nouns", "Ages", "N", "Nouns", "Ages","N")) %>%
  add_header_above(c("", "Comprehension" = 3, "Production" = 3)) %>%
  column_spec(1, bold = TRUE) %>%
  kable_styling(position = "center")

@@

Since the free association norms are available only in English, we used the hand-checked translation equivalents available in Wordbank, which allowed us to use the English association norms across 10 languages. Semantic associations are not necessarily shared across languages, but we use this technique as a reasonable first approximation. In support of this approximation, @youn2016 showed that semantic networks across languages share substantial similarities. Furthermore, using the same association data across languages does not necessarily mean that the resulting networks will grow in a similar fashion. For instance, the set of words acquired by children as well as the order of word acquisition can vary from language to language leading to possibility different learning strategies.

@@

In previous studies, two nodes were linked if they had an edit distance of 1 [@storkel2009; @carlson2014; @stella2017]. However, these studies reported a contribution of phonological connectivity to noun learning when networks were built using a rich adult vocabulary. Since the focus of the current study is on the mechanism of growth, the networks should be based on children's early vocabulary. The latter, however, contains very few noun pairs with an edit distance of 1. Thus, we increased the threshold from 1 to 2, that is, two nodes were related if their edit distance was equal to 1 or 2.\footnote{We also considered the case of an edit distance of 1 as well as the continuous measure, i.e., the inverse edit distance without threshold. In both cases, the results were weaker than those obtained with a threshold of 2. We did not consider the case of a threshold larger than 2 since many short pairs appear phonologically unrelated when the edit distance is 3 or more (e.g., "cat"/"dog").} The connectivity of a given node was characterized with its *degree*: the number of links it shares with other words.

@@

We start by analyzing word connectivity in the global (static) network. We constructed this network using nouns learned by the oldest age for which we have CDI data (e.g., in English this corresponds, in comprehension, to the network by 18 months, and in production, to the network by 30 months). This global network is the end-state towards which both INT and EXT converge by the last month of learning. Moreover, following @hills2009, we used this end-state network as a proxy for the external connectivity in the learning environment. Below we analyze properties of these global networks that may a priori hint at an INT- or EXT-like growth.

@@

```{r corrProd, out.width = "\\textwidth", fig.pos = "!h", fig.height=5, fig.cap = "Production data (Age of acquisition) as predicted by the degree (i.e., connectivity) in this network. Results are shown in each language for phonological and semantic networks. Each point is a word, with lines indicating linear model fits, and numbers indicating the Pearson correlation coefficients."}

@@

```{r corrComp, out.width = "\\textwidth", fig.pos = "!h", fig.height=5, fig.cap = "Comprehension data (Age of acquisition) as predicted by the degree (i.e., connectivity) in this network. Results are shown in each language for phonological and semantic networks. Each point is a word, with lines indicating linear model fits, and numbers indicating the Pearson correlation coefficients."}
corr_comp
#plot_grid(plot_correlation_prod , NULL, plot_correlation_comp , labels = c("A", "", "B"), ncol = 1, align = "v", rel_heights = c(1.3, 0.1, 1.3))

@@

Figures \@ref(fig:corrProd) and \@ref(fig:corrComp) show how the age of acquisition in production and comprehension, respectively, correlates with the degree (or indegree for the semantic networks). For ease of visual comparison, the predictor (i.e., the degree) was centered and scaled. The plots show, overall, a negative correlation between the month of acquisition and the degree.  In production data, the average correlation across languages was `r ps_ave` ($SD=$ `r ps_sd`) for the semantic networks and `r pp_ave` ($SD=$ `r pp_sd`) for the phonological networks.  In comprehension data, the average correlation was `r cs_ave` ($SD=$ `r cs_sd`) for the semantic networks and `r cp_ave` ($SD=$ `r cp_sd`) for the phonological networks. These results indicate that nouns with higher degrees are generally learned earlier, thus replicating previous findings in English [@storkel2009; @hills2009] and extending these findings to 10 different languages, generally, in both production- and comprehension-based vocabularies.

@@

We estimated the parameter $\beta$ using a Bayesian approach. The inference was performed using the probabilistic programming language  WebPPL [@dippl]. We defined a uniform prior over $\beta$, and at each month, we computed the likelihood function over words that could possibly enter the lexicon at that month, fit to the words that have been learned at that month (using Formula 1). Markov Chain Monte Carlo sampling resulted in a posterior distribution over $\beta$, which we summarized in Figure \@ref(fig:growthPred). The results replicate Hills et al.'s original finding regarding the semantic network in English and the production-based vocabulary, which is that this network grows by EXT, not by INT. Crucially, our results show that, generally speaking, this finding generalizes to comprehension-based vocabulary, and holds across languages. This generalization was obtained in both the semantic^[One could imagine that the fact of using English free association norms cross-linguistically would decrease the effect of non-English semantic networks because of possible cultural differences. However, our findings do not support this assumption, rather, it supports our initial approximation about the shared nature of the semantic similarity measure. That said, this approximation is not perfect. For example there is evidence that a small part of the variance in free association data can be explained by phonological similarity [@kachergis2011; @matusevych2018], thus leading to possibly minor cross-linguistic differences.] and phonological domains.

@@

Above we showed that the way semantic and phonological information is structured in the learning environment contributes to noun learning (via EXT) across languages. However, we know that other factors influence learning as well [e.g., @braginsky]. Next we investigated how semantic and phonological connectivity interact with two other factors. The first one is word frequency, a well studied factor shown to predict the age of acquisition in a reliable fashion [e.g. @goodman2008]. The second factor is word length, which was shown to correlate with phonological connectivity: Shorter words are more likely to have higher connectivity [@pisoni1985; @vitevitch2005].

@@

For word length, we counted the number of phonemes in our generated IPA transcription. For word frequency, we used the frequency estimates from @braginsky where unigram counts were derived based on CHILDES corpora in each language [@macwhinney2014]. Although these frequency counts use transcripts from independent sets of children, they are based on large samples, and this allows us to average out possible differences between children and the specificities of their input [see @goodman2008 for a similar research strategy].

@@

The findings for the new predictors were as follows. Overall, frequency is the largest and most consistent predictor of age of acquisition in both comprehension and production data and across languages, endorsing results for nouns across a variety of analyses [@goodman2008; @braginsky; @roy2015]. Word length is more predictive for production than comprehension (and this difference is very clear in the global model), replicating previous work [@braginsky]. Thus, word length seems to reflect the effects of production's constraints rather than comprehension's constraints, i.e., longer words are harder to articulate but they may not be significantly more difficult to store and access.

@@

We found that the original findings reported in @hills2009 generalize well across all these dimensions. First, just like semantic networks, phonological networks grow via the externally-driven scenario (EXT), not by the internally-driven mechanism (INT). Second, comprehension-based vocabularies grow in a way similar to production-based vocabularies. Finally, the findings were, overall, similar across the 10 languages we tested. Although we find some cross-linguistic variation when semantic and phonological networks were pitted against frequency and length, this variability is to be taken with a grain of salt as it might be exaggerated in our study by several factors such as the limited and partially-overlapping set of nouns for each language, measurement error due to the sample of acquisition data, the sample of frequency data, and the translation of association norms. In fact, both phonological and semantic connectivity are significant predictors above and beyond frequency and length when data are pooled across languages.

@@

For example, in the semantic domain, growth according to EXT could be explained by a mechanism similar to cross-situational learning [@smith2008; @mcmurray2012; @yurovsky2015]. According to this mechanism, children track the co-occurrence of concrete nouns with their possible semantic referents. The referent of a word heard in only one naming situation can be ambiguous (e.g., when the word "ball" is heard for the first time in the presence of both a ball and a chair), but hearing the same word in a diversity of semantic contexts allows the learner to narrow down the set of possible word-object mappings. In our case, free association (used to determine semantic network connectivity) is related to contextual co-occurrence [@griffiths07; @fourtassi2013], meaning that highly connected words will tend to occur in a variety of speech and referential contexts.  This fact makes such words easier to learn because they have more referential disambiguating cues across learning contexts. Crucially, children can learn these words without necessarily knowing the meaning of all other words with which they co-occur (hence the similarity with EXT). This possibility is supported by the finding that words' diversity of occurrence in child directed speech predicts their age of learning [@stella2017; @hills2010].

@@

This work shares a number of limitations with previous studies using similar research strategy and datasets. Chief among these limitations is the fact that the age of word acquisition is computed using different children at different ages (due to the fact that available CDI data is mainly cross-sectional). Such a measure has been shown to be valid and reliable [@fenson94], and has allowed researchers to study important aspects of word learning [@hills2009; @stella2017; @storkel2009; @goodman2008; @braginsky]. In our case, the use of cost-effective cross-sectional data has allowed us to leverage large-scale studies across several languages. That said, it is important to remember that this type of data can only inform us about the learning trajectory of the "average" child. Although our study endorses, overall, the externally-driven account of network growth, this does not mean individual children never use some variant of INT or some combination of both INT and EXT [@beckage]. To illustrate, some children develop "islands of expertise", that is, well organized knowledge about a certain topic (e.g., birds or dinosaurs). This prior knowledge enables these children to learn new related words more easily [e.g., @chi1983].
